---
title: "Data Science Midterm"
output: html_document
---

```{r Read Data}
library(tidyverse)
library(lubridate)

setwd("/Users/nancystetson/Google Drive/Midterm/Midterm Presentation")
rents <- read_csv("/Users/nancystetson/Google Drive/Midterm/CraigslistData/sfbay.csv")

dim(rents)

```


Next filter out listings without latitude and longitude
``` {r Filter}

rents <- rents %>% filter(!(is.na(latitude))) %>% filter(!(is.na(longitude)))
dim(rents)

```

Test out Google Directions API

```{r Google Maps API}
# API Key: AIzaSyAJL4gJxxAsXnHZ4sl5m1oYSv6GpFHCWI4

# Package option 1: Gmapsdistance
library(gmapsdistance)

# Set the key
key = "AIzaSyAJL4gJxxAsXnHZ4sl5m1oYSv6GpFHCWI4"
set.api.key("AIzaSyAJL4gJxxAsXnHZ4sl5m1oYSv6GpFHCWI4")

time <- ymd_hms("2016-11-04 08:00:08 PDT") # In november
sample_time <- as.numeric(time)

# Test directions output
x <- gmapsdistance(origin = "37.659645+-122.037708", # Random Location In Hayward Hills
                   destination = "37.784765+-122.407737",  #Downtown SF
                   mode = "driving",
                   traffic_model = "optimistic", # Can also be pessimistic
                   departure = sample_time) 
x

# Output is strange. Speculating it is in seconds and meters

x$Time / 60
# An hour and 3 minutes - google maps has it at hour and 6 right now

(x$Distance / 1000) * 0.621371
# Google maps has it at 30.6 miles

# Update: Traffic mode is only when dept time is set
# Also: best to use a range between pes. and opt., best guess is even more opt than opt
# SOLVED(Note - after some testing, it doesn't appear that when using dept 
# time it is taking into account traffic.  Duration is 40 min
# When leaving ten minutes from not
# But it is more than an hour otherwise )

```

```{r Other Package}

# Try out other API option
#library(googleway)


y <- google_distance(origins = "37.659645+-122.037708", # same locations as in the last example
               destinations = "37.784765+-122.407737",
                mode = "driving",
               units = "imperial",
               key = key,
               simplify = T)

# Don't like the format this is in, seems more sophisticated, but annoying to deal with
# use the other method until it doesn't work

```


# Next steps: Try to make the scatterplot of my dreams
``` {r plot it}
# First baby step, test this out for multiple distances
# Nope nevermind first coerce lat and lon into the right format

rents <- rents %>% mutate( location = paste(latitude, "+", longitude, sep = ""))
# Wasn't actually that hard

# Make a small data set
trunc_rents <- rents[1:100,]

x <- gmapsdistance(origin = trunc_rents$location, # 100 locations
                   destination = "37.784765+-122.407737",  #Downtown SF
                   mode = "driving",
                   traffic_model = "optimistic", # Can also be pessimistic
                   departure = sample_time) 
# took a long time

time_to_sf <- as.data.frame(x$Time)
time_to_sf$time_to <- time_to_sf$`Time.37.784765+-122.407737`
time_to_sf <- time_to_sf %>% 
              select(-`Time.37.784765+-122.407737`)

time_to_sf  <- time_to_sf %>%  mutate(minites = time_to /60) 

trunc_rents <- left_join(time_to_sf, trunc_rents, by = c("or" = "location"))


ggplot(trunc_rents, aes(minites, rent_sqft)) + geom_point() +
  geom_smooth() +
  xlab("Minutes to Downtown SF") + ylab("Rent Per SqFt" ) + ggtitle("Rent Versus Commute") +
  theme_minimal()

ggplot(trunc_rents, aes(minites, rent)) + geom_point() +
  geom_smooth() +
  xlab("Minutes to Downtown SF") + ylab("Rent" ) + ggtitle("Rent Versus Commute") +
  theme_minimal()

write_csv(trunc_rents, "ahundredrents_withdist.csv")

# Yay!
```


``` {r Map it}

library(leaflet)
library(rgeos)

map <-leaflet(data = trunc_rents1) %>% addProviderTiles("Stamen.TonerLite") %>%
  addCircles(radius = trunc_rents1$rent/10, popup = ~paste(round(minites)))
map

```


Ok next steps. I want to randomly sample 1000 listings (because Google has a limit). 
But I also want to continue to add to this day after day, so I can make my sample larger.
To do this, I am going to create an index the length of my data, and then jumble it, then attach
it to my data frame (rents). Then I can sort by this index and run the firs set today, the second set tomorrow, and so on. 

``` {r Create a Sample}
################################
# NOTE - I forgot to set a seed, so don't rerum this. will use written csv #
# first cleared workspace

#rents <- read_csv("/Users/nancystetson/Google Drive/Midterm/CraigslistData/sfbay.csv")
#rents <- rents %>% filter(!(is.na(latitude))) %>% filter(!(is.na(longitude)))
#dim(rents)
#rents <- rents %>% mutate( location = paste(latitude, "+", longitude, sep = ""))

# next create index
#index <- seq(1,length(rents$rent))

#index <- sample(index)

#rents$index <- index

#rents <- arrange(rents, index) 

#write_csv(rents, "rents_indexed.csv")
```


Next , for indices 1:1000, find the pessimistic and optimistic time to 
Downtown SF, arriving by 9 am on the day of our midterm.

``` {r}

thursday <- as.numeric(ymd_hms("2016-10-06 09:00:00 PDT"))

rents_to_1000 <- rents[1:1000,]

opt <- gmapsdistance(origin = rents_to_1000$location, # 100 locations
                   destination = "37.784765+-122.407737",  #Downtown SF
                   mode = "driving",
                   traffic_model = "optimistic", # Can also be pessimistic
                   arrival = thursday) 

ptm <- proc.time()
pes <- gmapsdistance(origin = rents_to_1000$location, # 100 locations
                   destination = "37.784765+-122.407737",  #Downtown SF
                   mode = "driving",
                   traffic_model = "pessimistic", # Can also be pessimistic
                   arrival = thursday) 
proc.time() - ptm


optimistic <- as.data.frame(opt$Time)
pessimistic <- as.data.frame(pes$Time)

optimistic$opt_time <- optimistic$`Time.37.784765+-122.407737`
optimistic$pes_time <- pessimistic$`Time.37.784765+-122.407737`

time_to <- optimistic
time_to <- time_to %>% 
              select(-`Time.37.784765+-122.407737`)

time_to  <- time_to %>% 
            mutate(opt_minutes = opt_time /60) %>%
            mutate(pes_minutes = pes_time / 60)

rents_to_1000 <- left_join(time_to, rents_to_1000, by = c("or" = "location"))


```


```{r}

rents_to_1000 <- read_csv('rents_to_1000.csv')

rbind(rents_to_1000, rents_to_2000)
